---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

#GRAM-NEGATIVE AND GRAM-POSITIVE BACTERIA CULTURE

##Activating Libraries
```{r}
library(readxl)
library(ggpubr)
library(forcats)
library(ggsci)
```

#Import table
```{r}
setwd("~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/GitHub/data/")
counts_CFU <- read.csv("metadata_full.csv")
counts_CFU$Treatment <- factor(counts_CFU$Treatment, levels = c("Control", "Antibiotic"))
```

#GRAM NEGATIVES

##Difference between day -1 and week 1 within the same treatment group
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_CFU.g` %notin% NA, study_ID %notin% "MA027", Time_tx %notin% "Week 11") %>%
  filter(Time_tx %in% c("Week 1", "Day -1")) %>% 
  dplyr::group_by(Treatment) %>%
  rstatix::wilcox_test(`GN_CFU.g` ~ Time_tx, alternative = "greater", paired = T) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```
##Difference between treatment groups by time point (day -1, week 1-5), PAIRED TEST 
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_CFU.g` %notin% NA, study_ID %notin% "MA027", Time_tx %notin% c("Week 11","Week 7","Week 9")) %>% #Filtering MA027 because it was drawn from the study in week 1. Not including weeks 7 and 9 because one animal was drawn from the study in week 7 (see chunk below) and the analysis could not be paired for those weeks. Not including Week 11 because we only took samples of 4 out of 20 animals in that week.
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GN_CFU.g` ~ Treatment, alternative = "greater", paired = T) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

##Difference between treatment groups by time point, weeks 7 and 9 PAIRED (1 cow was draw of the study on week7)
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_CFU.g` %notin% NA, study_ID %notin% c("MA027","MA028"), Time_tx %in% c("Week 7","Week 9")) %>% #Filtering MA028 because it was paired with MA029 which was drawn from the study on week 7
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GN_CFU.g` ~ Treatment, alternative = "greater", paired = T) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

##Significance over the time period (regardless of treatment)
```{r}
stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_CFU.g` %notin% NA) %>%
  dplyr::filter(Time_tx %notin% "Week 11") %>%
  kruskal_test(`GN_CFU.g` ~ Time_tx) 
stat.test
```

##Total Gram-negatives plot
```{r}
library(ggpubr)
total_line <- counts_CFU %>% 
  dplyr::filter(`GN_CFU.g` %notin% NA, study_ID %notin% "MA027", Time_tx %notin% "Week 11") %>%
  ggline(x = "Time_tx", y = "GN_CFU.g", color = "Treatment", add = c("mean_se"),  
         ylab = "Total Gram-negative CFU/g", xlab = "Time to treatment", palette = "jama") + 
  scale_y_continuous(labels = function(x) format(x, scientific = TRUE)) +  
  annotate("text", x = 1, y = 1.2e+06, label = expression("0.027"), colour = "black") +  
  annotate("text", x = 2, y = 1.8e+06, label = expression("0.006"), colour = "black") +  
  annotate("text", x = 3, y = 1.2e+06, label = expression("0.522"), colour = "black") +  
  annotate("text", x = 4, y = 1.5e+06, label = expression("0.435"), colour = "black") +  
  annotate("text", x = 5, y = 0.92e+06, label = expression("0.062"), colour = "black") +  
  annotate("text", x = 6, y = 1e+06, label = expression("0.198"), colour = "black") +  
  annotate("text", x = 7, y = 1.6e+06, label = expression("0.524"), colour = "black") +  
  annotate("text", x = 4, y = 2.5e+06, label = expression(paste("Kruskal-Wallis, ", paste(italic('p'))," = 0.037"))) +
  geom_signif(aes(y = `GN_CFU.g`, x = Time_tx),comparisons = list(c("Day -1","Week 1"),
                                                                 c("Week 1","Week 2"),
                                                                 c("Week 2","Week 3"),
                                                                 c("Week 3","Week 5"),
                                                                 c("Week 5","Week 7"), 
                                                                 c("Week 7","Week 9")), 
              test = "wilcox.test", y_position = c(1.8e+06,1.9e+06,1.5e+06,1.7e+06,1e+06,1.6e+06), color = "gray40", tip_length = 0.002, test.args=list(alternative = "less", var.equal = FALSE, paired=FALSE)) + 
  theme(axis.title.x = element_blank())

total_line
```
##Average Total Gram-negatives per week and treatment group
```{r}
counts_CFU %>% 
  dplyr::filter(`GN_CFU.g` %notin% NA, study_ID %notin% "MA027", Time_tx %notin% "Week 11") %>%
  group_by(Treatment, Time_tx) %>% 
  summarise(totaGN = mean(`GN_CFU.g`))
```

#Gram-negatives AMPICILLIN resistant CFU/g
##Significance over the time period (regardless of treatment)
```{r}
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_Amp_per` %notin% NA) %>%
  rstatix::kruskal_test(`GN_Amp_per` ~ Time_tx) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

##Difference between treatment groups by time point (day -1, week 1-5), PAIRED TEST 
```{r}
stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_Amp_CFU.g` %notin% NA) %>%
  dplyr::filter(`GN_Amp_CFU.g` %notin% NA, study_ID %notin% "MA027", Time_tx %notin% c("Week 11","Week 7","Week 9")) %>% #Filtering MA027 because it was drawn from the study in week 1. Not including weeks 7 and 9 because one animal was drawn from the study in week 7 (see chunk below) and the analysis could not be paired for those weeks. Not including Week 11 because we only took samples of 4 out of 20 animals in that week.
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GN_Amp_CFU.g` ~ Treatment, alternative = "greater", paired = T) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```
##Difference between treatment groups by time point (weeks 7-9), PAIRED TEST 
```{r}
stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_Amp_CFU.g` %notin% NA) %>%
  dplyr::filter(`GN_Amp_CFU.g` %notin% NA, study_ID %notin% c("MA027","MA028"), Time_tx %in% c("Week 7","Week 9")) %>% #Filtering MA028 because it was paired with MA029 which was drawn from the study on week 7
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GN_Amp_CFU.g` ~ Treatment, alternative = "greater", paired = T) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

##GN Amp CFU/g line plot
```{r}
amp_line <- counts_CFU %>% 
  dplyr::filter(`GN_Amp_CFU.g` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>% 
  dplyr::filter(Time_tx %notin% "Week 11") %>%
  ggline(x = "Time_tx", y = "GN_Amp_CFU.g", color = "Treatment", add = c("mean_se"),  ylab = "Ampicillin(R) Gram-negative CFU/g", xlab = "Time to treatment", palette = "jama") +  
  annotate("text", x = 1, y = 1.3e+05, label = expression("0.874"), colour = "black") +  
  annotate("text", x = 2, y = 5e+04, label = expression("0.460"), colour = "black") +  
  annotate("text", x = 3, y = 2.2e+05, label = expression("0.257"), colour = "black") +  
  annotate("text", x = 4, y = 6.4e+04, label = expression("0.167"), colour = "black") +  
  annotate("text", x = 5, y = 2e+04, label = expression("0.368"), colour = "black") +  
  annotate("text", x = 6, y = 2e+04, label = expression("0.381"), colour = "black") +  
  annotate("text", x = 7, y = 1.3e+05, label = expression("0.152"), colour = "black") +  
  annotate("text", x = 4, y = 3e+05, label = expression(paste("Kruskal-Wallis, ", paste(italic('p'))," = 7.73e-05"))) +  scale_y_continuous(labels = function(x) format(x, scientific = TRUE)) +
  geom_signif(aes(y = `GN_Amp_CFU.g`, x = Time_tx),comparisons = list(c("Day -1","Week 1"),
                                                                 c("Week 1","Week 2"),
                                                                 c("Week 2","Week 3"),
                                                                 c("Week 3","Week 5"),
                                                                 c("Week 5","Week 7"), 
                                                                 c("Week 7","Week 9")), 
              test = "wilcox.test", test.args=list(alternative = "less", var.equal = FALSE, paired=FALSE), color = "gray40", tip_length = 0.002, y_position = c(80000,1.5e+05,1.8e+05,10000,2e+02,60000)) + 
  theme(axis.title.x = element_blank())


amp_line 
```
#Gram-negatives AMPICILLIN resistant percentage (%)
Percentage was calculated over the total number of Gram-negative bacteria

##Significance over the time period (regardless of treatment)
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_Amp_per` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  rstatix::kruskal_test(`GN_Amp_per` ~ Time_tx) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

##Difference between treatment groups by time point (day -1, week 1-5), PAIRED TEST 
```{r}
stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_Amp_per` %notin% NA, study_ID %notin% "MA027", Time_tx %notin% c("Week 11","Week 7","Week 9")) %>% #Filtering MA027 because it was drawn from the study in week 1. Not including weeks 7 and 9 because one animal was drawn from the study in week 7 (see chunk below) and the analysis could not be paired for those weeks. Not including Week 11 because we only took samples of 4 out of 20 animals in that week.
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GN_Amp_per` ~ Treatment, alternative = "less", paired = T) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```
##Difference between treatment groups by time point (weeks 7-9), PAIRED TEST 
```{r}
stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_Amp_per` %notin% NA) %>%
  dplyr::filter(`GN_Amp_per` %notin% NA, study_ID %notin% c("MA027","MA028"), Time_tx %in% c("Week 7","Week 9")) %>% #Filtering MA028 because it was paired with MA029 which was drawn from the study on week 7
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GN_Amp_per` ~ Treatment, alternative = "less", paired = T) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

##GN Amp percentage line plot
```{r}
amp_line_per <- counts_CFU %>% 
  dplyr::filter(`GN_Amp_per` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>% 
  dplyr::filter(Time_tx %notin% "Week 11") %>%
  ggline(x = "Time_tx", y = "GN_Amp_per", color = "Treatment", add = c("mean_se"),  ylab = "Ampicillin(R) Gram-negative %", xlab = "Time to treatment", palette = "jama") +  
  annotate("text", x = 1, y = 13, label = expression("0.0026"), colour = "black") +  
  annotate("text", x = 2, y = 5, label = expression("0.245"), colour = "black") +  
  annotate("text", x = 3, y = 15, label = expression("0.785"), colour = "black") +  
  annotate("text", x = 4, y = 11, label = expression("0.702"), colour = "black") +  
  annotate("text", x = 5, y = 3, label = expression("0.185"), colour = "black") +  
  annotate("text", x = 6, y = 3, label = expression("0.234"), colour = "black") +  
  annotate("text", x = 7, y = 5, label = expression("0.516"), colour = "black") +  
  annotate("text", x = 4, y = 22, label = expression(paste("Kruskal-Wallis, ", paste(italic('p'))," = 8.72e-05"))) +
  geom_signif(aes(y = `GN_Amp_per`, x = Time_tx),comparisons = list(c("Day -1","Week 1"),
                                                                 c("Week 1","Week 2"),
                                                                 c("Week 2","Week 3"),
                                                                 c("Week 3","Week 5"),
                                                                 c("Week 5","Week 7"), 
                                                                 c("Week 7","Week 9")), 
              test = "wilcox.test", test.args=list(alternative  = "less", var.equal = FALSE, paired=FALSE), color = "gray40", tip_length = 0.002, y_position = c(9,11,13,7,.1,2)) + 
  theme(axis.title.x = element_blank())


amp_line_per
```

#Gram-negatives CEFTIOFUR resistant CFU/g
##Significance over the time period (regardless of treatment)##Significance over the time period (regardless of treatment)
```{r}
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_Cef_per` %notin% NA) %>%
  rstatix::kruskal_test(`GN_Cef_per` ~ Time_tx) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```
##Average by time point
```{r}
counts_CFU %>% 
  dplyr::filter(`GN_Cef_CFU.g` %notin% NA, study_ID %notin% "MA027") %>%
  group_by(Time_tx, Treatment) %>% 
  summarise(`Ceftiofur_CFU/g` = mean(`GN_Cef_CFU.g`))
```
##Difference between treatment groups by time point (day -1, week 1-5), PAIRED TEST 
```{r}
stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_Cef_CFU.g` %notin% NA, study_ID %notin% "MA027", Time_tx %notin% c("Week 11","Week 7","Week 9")) %>% #Filtering MA027 because it was drawn from the study in week 1. Not including weeks 7 and 9 because one animal was drawn from the study in week 7 (see chunk below) and the analysis could not be paired for those weeks. Not including Week 11 because we only took samples of 4 out of 20 animals in that week.
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GN_Cef_CFU.g` ~ Treatment, alternative = "less", paired = T) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```
##Difference between treatment groups by time point (weeks 7-9), PAIRED TEST 
```{r}
stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_Cef_CFU.g` %notin% NA) %>%
  dplyr::filter(`GN_Cef_CFU.g` %notin% NA, study_ID %notin% c("MA027","MA028"), Time_tx %in% c("Week 7","Week 9")) %>% #Filtering MA028 because it was paired with MA029 which was drawn from the study on week 7
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GN_Cef_CFU.g` ~ Treatment, alternative = "less", paired = T) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```
##Ceftiofur CFU/g plot
```{r}
cef_line <- counts_CFU %>% 
  dplyr::filter(`GN_Cef_CFU.g` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>% 
  dplyr::filter(Time_tx %notin% "Week 11") %>%
  ggline(x = "Time_tx", y = "GN_Cef_CFU.g", color = "Treatment", add = c("mean_se"),  ylab = "Ceftiofur(R) Gram-negative CFU/g", xlab = "Time to treatment", palette = "jama") +  
  annotate("text", x = 1, y = 360, label = expression("0.723"), colour = "black") +  
  annotate("text", x = 2, y = 380, label = expression("0.277"), colour = "black") +  
  annotate("text", x = 3, y = 620, label = expression("0.254"), colour = "black") +  
  annotate("text", x = 4, y = 100, label = expression("0.568"), colour = "black") +  
  annotate("text", x = 5, y = 100, label = expression("0.963"), colour = "black") +  
  annotate("text", x = 6, y = 150, label = expression("0.664"), colour = "black") +  
  annotate("text", x = 7, y = 300, label = expression("0.991"), colour = "black") +  
  annotate("text", x = 4, y = 1000, label = expression(paste("Kruskal-Wallis, ", paste(italic('p'))," = 0.0032"))) +
  geom_signif(aes(y = `GN_Cef_CFU.g`, x = Time_tx),comparisons = list(c("Day -1","Week 1"),
                                                                 c("Week 1","Week 2"),
                                                                 c("Week 2","Week 3"),
                                                                 c("Week 3","Week 5"),
                                                                 c("Week 5","Week 7"), 
                                                                 c("Week 7","Week 9")), 
              test = "wilcox.test", test.args=list(alternative = "less", var.equal = FALSE, paired=FALSE), color = "gray40", tip_length = 0.002, y_position = c(400,650,680,100,150,300)) + 
  theme(axis.title.x = element_blank())


cef_line 
```
#Gram-negatives CEFTIOFUR resistant percentage (%)
Percentage was calculated over the total number of Gram-negative bacteria
##Significance over the time period (regardless of treatment)
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_Cef_per` %notin% NA) %>%
  rstatix::kruskal_test(`GN_Cef_per` ~ Time_tx) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

##Difference between treatment groups by time point (day -1, week 1-5), PAIRED TEST 
```{r}
stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_Cef_per` %notin% NA, study_ID %notin% "MA027", Time_tx %notin% c("Week 11","Week 7","Week 9")) %>% #Filtering MA027 because it was drawn from the study in week 1. Not including weeks 7 and 9 because one animal was drawn from the study in week 7 (see chunk below) and the analysis could not be paired for those weeks. Not including Week 11 because we only took samples of 4 out of 20 animals in that week.
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GN_Cef_per` ~ Treatment, alternative = "less", paired = T) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```
##Difference between treatment groups by time point (weeks 7-9), PAIRED TEST 
```{r}
stat.test <- counts_CFU %>% 
  dplyr::filter(`GN_Cef_per` %notin% NA) %>%
  dplyr::filter(`GN_Cef_per` %notin% NA, study_ID %notin% c("MA027","MA028"), Time_tx %in% c("Week 7","Week 9")) %>% #Filtering MA028 because it was paired with MA029 which was drawn from the study on week 7
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GN_Cef_per` ~ Treatment, alternative = "less", paired = T) %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

##GN Ceftiofur resistant percentage line plot
```{r}
cef_line_per <- counts_CFU %>% 
  dplyr::filter(`GN_Cef_per` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>% 
  dplyr::filter(Time_tx %notin% "Week 11") %>%
  ggline(x = "Time_tx", y = "GN_Cef_per", color = "Treatment", add = c("mean_se"),  ylab = "Ceftiofur(R) Gram-negative %", xlab = "Time to treatment", palette = "jama") +  
  annotate("text", x = 1, y = .08, label = expression("0.534"), colour = "black") +  
  annotate("text", x = 2, y = .12, label = expression("0.336"), colour = "black") +  
  annotate("text", x = 3, y = .07, label = expression("0.290"), colour = "black") +  
  annotate("text", x = 4, y = .03, label = expression("0.807"), colour = "black") +  
  annotate("text", x = 5, y = .03, label = expression("0.977"), colour = "black") +  
  annotate("text", x = 6, y = .04, label = expression("0.135"), colour = "black") +  
  annotate("text", x = 7, y = .07, label = expression("0.838"), colour = "black") +  
  annotate("text", x = 4, y = .20, label = expression(paste("Kruskal-Wallis, ", paste(italic('p'))," = 0.002"))) +
  geom_signif(aes(y = `GN_Cef_per`, x = Time_tx),comparisons = list(c("Day -1","Week 1"),
                                                                 c("Week 1","Week 2"),
                                                                 c("Week 2","Week 3"),
                                                                 c("Week 3","Week 5"),
                                                                 c("Week 5","Week 7"), 
                                                                 c("Week 7","Week 9")), 
              test = "wilcox.test", test.args=list(alternative = "less", var.equal = FALSE, paired=FALSE), color = "gray40", tip_length = 0.002, y_position = c(.10,.12,.07,.02,.03,.05)) + 
  theme(axis.title.x = element_blank())


cef_line_per
```

#GRAM POSITIVES total

#Average Gram-positives by treatment and time point
```{r}
counts_CFU %>% 
  dplyr::filter(`GP_CFU.g` %notin% NA, study_ID %notin% "MA027", Time_tx %notin% "Week 11") %>%
  group_by(Treatment, Time_tx) %>% 
  summarise(totaGN = mean(`GP_CFU.g`))
```

##Difference between time-points per treatment group
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GP_CFU.g` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  dplyr::group_by(Treatment) %>%
  rstatix::wilcox_test(`GP_CFU.g` ~ Time_tx, alternative = "g", paired = T) %>%
    rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p")
stat.test
```

##Difference between treatments by time point
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GP_CFU.g` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  dplyr::filter(Time_tx %in% c("Day -1", "Week 1")) %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GP_CFU.g` ~ Treatment, alternative = "greater", paired = T) %>%
    rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p") %>%
  add_xy_position(x = "Time_tx", dodge = 0.8)
stat.test
```

##Total Gram-positives plot
```{r}
library(ggpubr)
library(ggsci)
GPtotal_boxplot = counts_CFU %>% 
  dplyr::filter(`GP_CFU.g` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  ggboxplot(x = "Time_tx", y = "GP_CFU.g", color = "Treatment", palette = "jama", 
            fill = "Treatment", add = c("jitter"), notch = F, outlier.shape = NA) +
  labs(x = "Time to treatment", y = "Total Gram-positive CFU/g") + 
  scale_fill_jama(alpha = 0.5) + 
  theme(legend.position="right") + 
  scale_y_continuous(labels = function(x) format(x, scientific = TRUE)) + 
  stat_pvalue_manual(stat.test,  label = "p", tip.length = 0) +  
  geom_signif(aes(y = `GP_CFU.g`, x = Time_tx),
              comparisons = list(c("Day -1","Week 1")), 
              test = "wilcox.test", test.args=list(alternative = "less", var.equal = FALSE, paired=FALSE), color = "gray40", tip_length = 0.002, y_position = c(8.5e+06)) + 
  theme(axis.title.x = element_blank())

GPtotal_boxplot
```
#GRAM POSITIVES Ceftiofur resistant CFU/g
##Average Gram-positives resistant to ceftiofur by treatment and time point
```{r}
counts_CFU %>% 
  dplyr::filter(`GP_Cef_CFU.g` %notin% NA, study_ID %notin% "MA027", Time_tx %notin% "Week 11") %>%
  group_by(Treatment, Time_tx) %>% 
  summarise(totaGN = mean(`GP_Cef_CFU.g`))
```

##Difference between time-points per treatment group
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GP_Cef_CFU.g` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  dplyr::group_by(Treatment) %>%
  rstatix::wilcox_test(`GP_Cef_CFU.g` ~ Time_tx, alternative = "g", paired = T) %>%
    rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p")
stat.test
```

##Difference between treatments by time point
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GP_Cef_CFU.g` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  dplyr::filter(Time_tx %in% c("Day -1", "Week 1")) %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GP_Cef_CFU.g` ~ Treatment, alternative = "greater", paired = T) %>%
    rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p") %>%
  add_xy_position(x = "Time_tx", dodge = 0.8)
stat.test
```

##Boxplot
```{r}
library(ggpubr)
library(ggsci)
GPcef_boxplot = counts_CFU %>% 
  dplyr::filter(`GP_Cef_CFU.g` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  ggboxplot(x = "Time_tx", y = "GP_Cef_CFU.g", color = "Treatment", palette = "jama", 
            fill = "Treatment", add = c("jitter"), notch = F, outlier.shape = NA) +
  labs(x = "Time to treatment", y = "Ceftiofur(R) Gram-positive CFU/g") + 
  scale_fill_jama(alpha = 0.5) + theme(legend.position="right") + 
  scale_y_continuous(labels = function(x) format(x, scientific = TRUE)) + 
  theme(axis.title.x = element_blank()) + 
  stat_pvalue_manual(stat.test,  label = "p", tip.length = 0) +  
  geom_signif(aes(y = `GP_Cef_CFU.g`, x = Time_tx),
              comparisons = list(c("Day -1","Week 1")), 
              test = "wilcox.test", test.args=list(alternative = "less", var.equal = FALSE, paired=FALSE), color = "gray40", tip_length = 0.002, y_position = c(8e+05))

GPcef_boxplot
```

#GP ceftiofur percentage
Percentage calculated over the total of Gram-positive per sample

##Average Gram-positives resistant to ceftiofur by treatment and time point
```{r}
counts_CFU %>% 
  dplyr::filter(`GP_Cef_per` %notin% NA, study_ID %notin% "MA027", Time_tx %notin% "Week 11") %>%
  group_by(Treatment, Time_tx) %>% 
  summarise(totaGN = mean(`GP_Cef_per`))
```

##Difference between time-points per treatment group
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GP_Cef_per` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  dplyr::group_by(Treatment) %>%
  rstatix::wilcox_test(`GP_Cef_per` ~ Time_tx, alternative = "g", paired = T) %>%
    rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p")
stat.test
```

##Difference between treatments by time point
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GP_Cef_per` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  dplyr::filter(Time_tx %in% c("Day -1", "Week 1")) %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GP_Cef_per` ~ Treatment, alternative = "less", paired = T) %>%
    rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p") %>%
  add_xy_position(x = "Time_tx", dodge = 0.8)
stat.test
```
##Boxplot
```{r}
library(ggpubr)
library(ggsci)
GPcef_boxplot_per = counts_CFU %>% 
  dplyr::filter(`GP_Cef_per` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  ggboxplot(x = "Time_tx", y = "GP_Cef_per", color = "Treatment", palette = "jama", 
            fill = "Treatment", add = c("jitter"), notch = F, outlier.shape = NA) +
  labs(x = "Time to treatment", y = "Ceftiofur(R) Gram-positive %") + 
  scale_fill_jama(alpha = 0.5) + 
  theme(legend.position="right") + 
  theme(axis.title.x = element_blank()) + 
  stat_pvalue_manual(stat.test,  label = "p", tip.length = 0) +  
  geom_signif(aes(y = `GP_Cef_per`, x = Time_tx),
              comparisons = list(c("Day -1","Week 1")), 
              test = "wilcox.test", test.args=list(alternative = "less", var.equal = FALSE, paired=FALSE), 
              color = "gray40", tip_length = 0.002, y_position = c(90))

GPcef_boxplot_per
```
#GP Ampicillin CFU/g

##Average Gram-positives resistant to ampicillin by treatment and time point
```{r}
counts_CFU %>% 
  dplyr::filter(`GP_Amp_CFU.g` %notin% NA, study_ID %notin% "MA027", Time_tx %notin% "Week 11") %>%
  group_by(Treatment, Time_tx) %>% 
  summarise(totaGN = mean(`GP_Amp_CFU.g`))
```

##Difference between time-points per treatment group
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GP_Amp_CFU.g` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  dplyr::group_by(Treatment) %>%
  rstatix::wilcox_test(`GP_Amp_CFU.g` ~ Time_tx, alternative = "g", paired = T) %>%
    rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p")
stat.test
```

##Difference between treatments by time point
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GP_Amp_CFU.g` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  dplyr::filter(Time_tx %in% c("Day -1", "Week 1")) %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GP_Amp_CFU.g` ~ Treatment, alternative = "less", paired = T) %>%
    rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p") %>%
  add_xy_position(x = "Time_tx", dodge = 0.8)
stat.test
```


##Boxplot
```{r}
library(ggpubr)
library(ggsci)
GPamp_boxplot = counts_CFU %>% 
  dplyr::filter(`GP_Amp_CFU.g` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  ggboxplot(x = "Time_tx", y = "GP_Amp_CFU.g", color = "Treatment", palette = "jama", 
            fill = "Treatment", add = c("jitter"), notch = F, outlier.shape = NA) +
  labs(x = "Time to treatment", y = "Ampicillin(R) Gram-positive CFU/g") + 
  scale_fill_jama(alpha = 0.5) + theme(legend.position="right") + 
  scale_y_continuous(labels = function(x) format(x, scientific = TRUE)) + 
  theme(axis.title.x = element_blank()) + 
  stat_pvalue_manual(stat.test,  label = "p", tip.length = 0) +  
  geom_signif(aes(y = `GP_Amp_CFU.g`, x = Time_tx),
              comparisons = list(c("Day -1","Week 1")), 
              test = "wilcox.test", test.args=list(alternative = "less", var.equal = FALSE, paired=FALSE), 
              color = "gray40", tip_length = 0.002, y_position = c(1.8e+05))

GPamp_boxplot
```

#GP Ampicillin percentage
Percentage calculated over the total of Gram-positive per sample

##Average Gram-positives resistant to ampicillin by treatment and time point
```{r}
counts_CFU %>% 
  dplyr::filter(`GP_Amp_per` %notin% NA, study_ID %notin% "MA027", Time_tx %notin% "Week 11") %>%
  group_by(Treatment, Time_tx) %>% 
  summarise(totaGN = mean(`GP_Amp_per`))
```

##Difference between time-points per treatment group
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GP_Amp_per` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  dplyr::group_by(Treatment) %>%
  rstatix::wilcox_test(`GP_Amp_per` ~ Time_tx, alternative = "g", paired = T) %>%
    rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p")
stat.test
```

##Difference between treatments by time point
```{r}
library(dplyr)
library(rstatix)
`%notin%` <- Negate(`%in%`)

stat.test <- counts_CFU %>% 
  dplyr::filter(`GP_Amp_per` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  dplyr::filter(Time_tx %in% c("Day -1", "Week 1")) %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(`GP_Amp_per` ~ Treatment, alternative = "less", paired = T) %>%
    rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p") %>%
  add_xy_position(x = "Time_tx", dodge = 0.8)
stat.test
```
##Gram-positives resistant to ampicillin percentage plot
```{r}
library(ggpubr)
library(ggsci)
GPamp_boxplot_per = counts_CFU %>% 
  dplyr::filter(`GP_Amp_per` %notin% NA) %>%
  dplyr::filter(study_ID %notin% "MA027") %>%
  ggboxplot(x = "Time_tx", y = "GP_Amp_per", color = "Treatment", palette = "jama", 
            fill = "Treatment", add = c("jitter"), notch = F, outlier.shape = NA) +
  labs(x = "Time to treatment", y = "Ampicillin(R) Gram-positive %") + 
  scale_fill_jama(alpha = 0.5) + theme(legend.position="right") + 
  theme(axis.title.x = element_blank()) + 
  stat_pvalue_manual(stat.test,  label = "p", tip.length = 0) +  
  geom_signif(aes(y = `GP_Amp_per`, x = Time_tx),
              comparisons = list(c("Day -1","Week 1")), 
              test = "wilcox.test", test.args=list(alternative = "less", var.equal = FALSE, paired=FALSE), 
              color = "gray40", tip_length = 0.002, y_position = c(40))

GPamp_boxplot_per
```

```{r fig.height=10, fig.width=9}
culture_plots <- ggarrange(total_line, GPtotal_boxplot, cef_line, GPcef_boxplot, amp_line, GPamp_boxplot, nrow = 3,ncol = 2, common.legend = T, labels = c("A","B","C","D","E","F"), widths = c(1,.5,1,.5,1,.5))
culture_plots
```

```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Mastitis_project/figures/culture")
ggsave("culture_plots_KW.png", plot = culture_plots, width = 9, height = 12)
```

```{r fig.height=10, fig.width=9}
culture_plots_per <- ggarrange(cef_line_per, GPcef_boxplot_per, amp_line_per, GPamp_boxplot_per, nrow = 2,ncol = 2, common.legend = T, labels = c("A","B","C","D"), widths = c(1,.5,1,.5))
culture_plots_per
```

```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Mastitis_project/figures/culture")
ggsave("culture_plots_res_per.png", plot = culture_plots_per, width = 9, height = 9)
```

#Comparison percentage of Gram-positive resistant to ceftiofur vs. ampicillin within samples
```{r}
wilcox.test(counts_CFU$GP_Cef_per, counts_CFU$GP_Amp_per)
```

#Comparison percentage of Gram-negative resistant to ceftiofur vs. ampicillin within samples
```{r}
wilcox.test(counts_CFU$GN_Cef_per, counts_CFU$GN_Amp_per)
```

#Comparison prevalence Gram-positive ceftiofur resistant between treatments
```{r}
antibiotic <- counts_CFU %>% filter(Treatment %in% "Antibiotic",  Time_tx %in% c("Day -1", "Week 1"))
control <- counts_CFU %>% filter(Treatment %in% "Control", Time_tx %in% c("Day -1", "Week 1"))

prevalence_GPcef_antibiotic <- ifelse(antibiotic$GP_Cef_per>0,1,0)
prevalence_GPcef_control <- ifelse(control$GP_Cef_per>0,1,0)

wilcox.test(prevalence_GPcef_antibiotic, prevalence_GPcef_control)
```
#Comparison prevalence Gram-positive ampicillin resistant between treatments
```{r}
prevalence_GPamp_antibiotic <- ifelse(antibiotic$GP_Amp_per>0,1,0)
#prevalence_GPamp_antibiotic[is.na(prevalence_GPamp_antibiotic)] = 0

prevalence_GPamp_control <- ifelse(control$GP_Amp_per>0,1,0)
#prevalence_GPamp_control[is.na(prevalence_GPamp_control)] = 0

wilcox.test(prevalence_GPamp_antibiotic, prevalence_GPamp_control)
```

#Comparison prevalence Gram-negative ceftiofur resistant between treatments
```{r}
antibiotic_GN <- counts_CFU %>% filter(Treatment %in% "Antibiotic",  Time_tx %notin% c("Week 11"))
control_GN <- counts_CFU %>% filter(Treatment %in% "Control", Time_tx %notin% c("Week 11"))

prevalence_GNcef_antibiotic <- ifelse(antibiotic_GN$GN_Cef_per>0,1,0)
prevalence_GNcef_antibiotic[is.na(prevalence_GNcef_antibiotic)] = 0

prevalence_GNcef_control <- ifelse(control_GN$GN_Cef_per>0,1,0)
prevalence_GNcef_control[is.na(prevalence_GNcef_control)] = 0

wilcox.test(prevalence_GNcef_antibiotic, prevalence_GNcef_control)
```
#Comparison prevalence Gram-negative ampicillin resistant between treatments
```{r}
prevalence_GNamp_antibiotic <- ifelse(antibiotic_GN$GN_Amp_per>0,1,0)
prevalence_GNamp_antibiotic[is.na(prevalence_GNamp_antibiotic)] = 0

prevalence_GNamp_control <- ifelse(control_GN$GN_Amp_per>0,1,0)
prevalence_GNamp_control[is.na(prevalence_GNamp_control)] = 0

wilcox.test(prevalence_GNamp_antibiotic, prevalence_GNamp_control)
```
#Comparison variation of CFU/g Gram-negative resistant to ampicillin
```{r}
GN_counts <- counts_CFU %>% dplyr::filter(study_ID %notin% "MA027", Time_tx %notin% "Week 11")
kruskal.test(GN_counts$GN_Amp_CFU.g ~ GN_counts$Time_tx)
```

#Comparison variation of CFU/g Gram-negative resistant to ceftiofur
```{r}
kruskal.test(GN_counts$GN_Cef_CFU.g ~ GN_counts$Time_tx)
```

#Comparison variation of % Gram-negative resistant to ceftiofur
```{r}
kruskal.test(GN_counts$GN_Cef_per ~ GN_counts$Time_tx)
```

#Comparison variation of % Gram-negative resistant to ampicillin
```{r}
kruskal.test(GN_counts$GN_Amp_per ~ GN_counts$Time_tx)
```

#CEFTIOFUR resistant isolates
```{r}
#Import tables
setwd("~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/GitHub/data/")
cef_strains <- read.csv("ceftiofur_resistant_isolates.csv")
cef_strains$Treatment <- factor(cef_strains$Treatment, levels = c("Control", "Antibiotic"))
```

#Number of isolates per treatment
```{r}
cef_strains %>% 
  group_by(Treatment) %>%
  summarise(total_isolates=sum(Isolates))
```
#Plot showing the number of isolates per sampling point and treatment group
```{r}
cef_strains$Time_tx <- factor(cef_strains$Time_tx, levels = c("Day -1", "Week 1", "Week 2", "Week 3", "Week 5", "Week 7", "Week 9", "Week 11"))

cef_strains_number <-cef_strains %>% 
  group_by(Time_tx, Treatment) %>%
  summarise(total_isolates=sum(Isolates)) %>% 
  ggbarplot(x = "Time_tx", y = "total_isolates", color = "Treatment", fill="Treatment",  
            lab.pos = "in", label = T, lab.col = "black", 
            ylab = "Number of isolates", 
            xlab = "Time to treatment", 
            palette = "jama") 
cef_strains_number
```

