---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

#Import tables with network stats calculated with Gephi
```{r}
setwd("~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/GitHub/data/networks/stats")
day1_ab <- read.csv("day-1_antibiotic.csv")
day1_ct <- read.csv("day-1_control.csv")
week1_ab <- read.csv("week1_antibiotic.csv")
week1_ct <- read.csv("week1_control.csv")
week5_ab <- read.csv("week5_antibiotic.csv")
week5_ct <- read.csv("week5_control.csv")
week9_ab <- read.csv("week9_antibiotic.csv")
week9_ct <- read.csv("week9_control.csv")
```

#Adding atributes to each table before merging them
```{r}
library(dplyr)
day1_ab <- day1_ab %>% mutate(Time_tx = "Day -1", Treatment = "Antibiotic")
day1_ct <- day1_ct %>% mutate(Time_tx = "Day -1", Treatment = "Control")
week1_ab <- week1_ab %>% mutate(Time_tx = "Week 1", Treatment = "Antibiotic")
week1_ct <- week1_ct %>% mutate(Time_tx = "Week 1", Treatment = "Control")
week5_ab <- week5_ab %>% mutate(Time_tx = "Week 5", Treatment = "Antibiotic")
week5_ct <- week5_ct %>% mutate(Time_tx = "Week 5", Treatment = "Control")
week9_ab <- week9_ab %>% mutate(Time_tx = "Week 9", Treatment = "Antibiotic")
week9_ct <- week9_ct %>% mutate(Time_tx = "Week 9", Treatment = "Control")
```

#joining tables into one dataframe
```{r}
network_stats <- day1_ab %>% 
  full_join(day1_ct) %>% 
  full_join(week1_ab) %>% 
  full_join(week1_ct) %>% 
  full_join(week5_ab) %>% 
  full_join(week5_ct) %>% 
  full_join(week9_ab) %>% 
  full_join(week9_ct)
```

#Making a new table only containing beta-lactam ARGs
```{r}
betalactams <- network_stats %>% 
  filter(class %in% "Beta-lactam")
betalactams
```

#DEGREE OF CENTRALITY
#Calculating the sum of degree centrality per group and sampling poing
```{r}
betalactams %>%
  dplyr::group_by(Time_tx, Treatment) %>%
  summarise(degree_sum = sum(Degree))
```

##Stats of degree between treatment groups for each sampling point
```{r}
library(rstatix)
stat.test <- betalactams %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(Degree ~ Treatment, alternative = "l") %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```
##Stats of degree between sampling points by treatment group
```{r}
library(rstatix)
stat.test <- betalactams %>%
  dplyr::group_by(Treatment) %>%
  rstatix::wilcox_test(Degree ~ Time_tx, alternative = "g") %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```


##plot
```{r}
library(ggpubr)
degree_line <- betalactams %>% 
  ggline(x = "Time_tx", y = "Degree", color = "gene", add = c("mean_se"), facet.by = "Treatment", palette = get_palette("simpsons",20), ylab = "Degree of centrality", xlab = "Time to treatment") 
degree_line
```
##Stats
```{r}
betalactams %>% 
  rstatix::kruskal_test(Degree ~ Time_tx)
```

```{r}
betalactams %>% 
  rstatix::kruskal_test(betweenesscentrality ~ Time_tx)
```

```{r}
library(rstatix)
stat.test <- betalactams %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(Degree ~ Treatment, alternative = "g") %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

```{r}
library(ggpubr)
degree_line <- betalactams %>% 
  ggline(x = "Time_tx", y = "Degree", color = "Treatment", add = c("mean_se"), palette = "jama", ylab = "Degree of centrality", xlab = "Time to treatment", remove = "point") +  
  annotate("text", x = 1, y = 18, label = expression("0.043"), colour = "black") +  
  annotate("text", x = 2, y = 12, label = expression("0.555"), colour = "black") +  
  annotate("text", x = 3, y = 10, label = expression("0.860"), colour = "black") +  
  annotate("text", x = 4, y = 35, label = expression("0.818"), colour = "black") +
    annotate("text", x = 2.5, y = 45, label = expression(paste("Kruskal-Wallis, ", paste(italic('p'))," = 0.261"))) +
  stat_compare_means(comparisons=list(c("Day -1","Week 1"), 
                                           c("Week 1","Week 5"),
                                           c("Week 5","Week 9")),
                     label.y=c(15, 17, 32),
                     tip.length = 0.002, 
                     method="wilcox.test", label="p.format", color="gray40",
                     method.args=list(alternative = "two.sided", var.equal = FALSE, paired=FALSE))

degree_line
```

#BETWEENESS CENTRALITY
##Stats
```{r}
library(rstatix)
stat.test <- betalactams %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(betweenesscentrality ~ Treatment, alternative = "l") %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

##Plot
```{r}
betcen_line <- betalactams %>% 
  ggline(x = "Time_tx", y = "betweenesscentrality", color = "gene", add = c("mean_se"), facet.by = "Treatment", palette = get_palette("simpsons",20), ylab = "Betweenness centrality", xlab = "Time to treatment") + labs(fill="Gene", color="Gene") + theme(legend.text = element_text(face = "italic"))
betcen_line
```

##Plot

```{r}
betalactams$Treatment <- factor(betalactams$Treatment, levels = c("Control", "Antibiotic"))
betcen_line <- betalactams %>% 
  ggline(x = "Time_tx", y = "betweenesscentrality", color = "Treatment", add = c("mean_se"), palette = "jama", ylab = "Betweenness centrality", xlab = "Time to treatment") + labs(fill="Gene", color="Gene") +  
  annotate("text", x = 1, y = 260, label = expression("0.008"), colour = "black") +  
  annotate("text", x = 2, y = 230, label = expression("0.725"), colour = "black") +  
  annotate("text", x = 3, y = 100, label = expression("0.950"), colour = "black") +  
  annotate("text", x = 4, y = 180, label = expression("0.787"), colour = "black") +
      annotate("text", x = 2.5, y = 350, label = expression(paste("Kruskal-Wallis, ", paste(italic('p'))," = 0.052"))) +
  stat_compare_means(comparisons=list(c("Day -1","Week 1"), 
                                           c("Week 1","Week 5"),
                                           c("Week 5","Week 9")),
                     label.y=c(250, 200, 150),
                     tip.length = 0.002, 
                     method="wilcox.test", label="p.format", color="gray40",
                     method.args=list(alternative = "two.sided", var.equal = FALSE, paired=FALSE))
betcen_line
```
```{r}
stat.test <- betalactams %>%
  dplyr::group_by(Treatment) %>%
  rstatix::wilcox_test(betweenesscentrality ~ Time_tx, alternative = "l") %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

```{r}
network_stats_plot <- ggarrange(degree_line, betcen_line, nrow = 1, labels = c("A", "B"), common.legend = T)
network_stats_plot
```

```{r}
setwd("~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/figures/networks/")
ggsave(plot=network_stats_plot, "centrality_bla-genes.png", width = 8, height = 4)
```

#CLUSTERING
##Stats
```{r}
stat.test <- betalactams %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(clustering ~ Treatment, alternative = "l") %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

##plot
```{r}
clustering_line <- betalactams %>% 
  ggline(x = "Time_tx", y = "clustering", color = "gene", add = c("mean_se"), facet.by = "Treatment", palette = get_palette("rickandmorty",20), ylab = "Clustering", xlab = "Time to treatment") 
clustering_line
```

#CLOSENESS CENTRALITY
##Stats
```{r}
stat.test <- betalactams %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(closnesscentrality ~ Treatment, alternative = "l") %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

##plot
```{r}
closness_line <- betalactams %>% 
  ggline(x = "Time_tx", y = "closnesscentrality", color = "gene", add = c("mean_se"), facet.by = "Treatment", palette = get_palette("rickandmorty",20), ylab = "Closeness Centrality", xlab = "Time to treatment") 
closness_line
```


#WEIGHTED DEGREE
##Stats
```{r}
stat.test <- betalactams %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(Weighted.Degree ~ Treatment, alternative = "l") %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

##plot
```{r}
wdegree_line <- betalactams %>% 
  ggline(x = "Time_tx", y = "Weighted.Degree", color = "gene", add = c("mean_se"), facet.by = "Treatment", palette = get_palette("rickandmorty",20), ylab = "Weighted Degree", xlab = "Time to treatment") 
wdegree_line
```

#EIGEN CENTRALITY
##Stats
```{r}
stat.test <- betalactams %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(eigencentrality ~ Treatment, alternative = "l") %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

##plot
```{r}
eigen_line <- betalactams %>% 
  ggline(x = "Time_tx", y = "eigencentrality", color = "gene", add = c("mean_se"), facet.by = "Treatment", palette = get_palette("rickandmorty",20), ylab = "Eigen Centrality", xlab = "Time to treatment") 
eigen_line
```

```{r}
genes_degree <- betalactams %>%
  dplyr::group_by(Time_tx, Treatment, gene) %>%
  summarise(degree_sum = sum(Degree))

genes_degree
```

```{r}
genes_betcen <- betalactams %>%
  dplyr::group_by(Time_tx, Treatment, gene) %>%
  summarise(betcen = sum(betweenesscentrality))

genes_betcen
```

```{r}
genes_betcen <- betalactams %>%
  dplyr::group_by(gene) %>%
  summarise(avg.betcen = mean(betweenesscentrality))

genes_betcen
```

#General network stats
```{r}
stat.test <- network_stats %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(Degree ~ Treatment, alternative = "l") %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

```{r}
stat.test <- network_stats %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(Weighted.Degree ~ Treatment, alternative = "l") %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

```{r}
stat.test <- network_stats %>%
  dplyr::group_by(Time_tx) %>%
  rstatix::wilcox_test(betweenesscentrality ~ Treatment, alternative = "l") %>%
 rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
stat.test
```

```{r}
`%notin%` <- Negate(`%in%`)
betalactams %>%
  filter(gene %notin% c("blaAmpC", "blaAmpH", "pbp2", "pbp4"), Time_tx %in% "Week 9") %>%
  group_by(Treatment) %>% 
  summarise(degree_mean = mean(Degree))
```
```{r}
betalactams %>%
  filter(Time_tx %in% "Week 9") %>%
  group_by(Treatment) %>% 
  summarise(degree_mean = mean(Degree))
```
```{r}
b9a <- betalactams %>% filter(Time_tx %in% "Week 9", Treatment %in% "Antibiotic") %>%  filter(gene %notin% c("blaAmpC", "blaAmpH", "pbp2", "pbp4"))
  
b9c <- betalactams %>% filter(Time_tx %in% "Week 9", Treatment %in% "Control") %>%  filter(gene %notin% c("blaAmpC", "blaAmpH", "pbp2", "pbp4"))
  
wilcox.test(b9a$Degree, b9c$Degree, alternative = "greater", paired = F)
```

```{r}
betalactams %>%
  dplyr::filter(Time_tx %in% "Week 9") %>%
  rstatix::wilcox_test(Degree ~ Treatment, alternative = "g") %>%
  rstatix::adjust_pvalue(method = "none") %>%
  rstatix::add_significance("p.adj")
```

