---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
#BACTERIA DIVERSITY AND COMPOSITION
##Activating libraries
```{r}
library(tidyverse)
library("ggsci")
library(ggpubr)
library(rstatix)
library(scales)
```

#Importing tables
```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Mastitis_project/GitHub/data/")

#Metadata
metadata =read.table("metadata.csv", sep = ",", header = TRUE,  na.strings = "") %>% 
   filter(Time_tx %in% c("Day -1","Week 1","Week 5", "Week 9"))
metadata$days_tx <- as.numeric(as.character(metadata$days_tx))
metadata$Treatment <- factor(metadata$Treatment, levels= c("Control","Antibiotic"))

#Alpha diversity (shannon, simpson, chao1) calculated through the COSMOS-ID pipeline
alpha =read.table("alpha_bacteria_abs_COSMOS-ID.tsv", sep = "\t", header = TRUE,  na.strings = "") 

#Beta diversity (Jaccard & Bray-curtis dissimilarity) calculated through the COSMOS-ID pipeline
beta = read.table("betapca_bacteria_abs_COSMOS-ID.tsv", sep = "\t", header = TRUE,  na.strings = "") 

#ARGs abundance score calculated through the COSMOS-ID pipeline
abundance = read.table("abundance_bacteria_abs_COSMOS-ID.tsv", sep = "\t", header = TRUE,  na.strings = "", check.names = F)
```

#Merging diversity tables with metadata
```{r}
alpha_met <- merge(metadata,alpha,by="Name")
alpha_beta_met <- merge(alpha_met,beta,by="Name")
```

#ALPHA DIVERSITY
##Shannon
```{r}
library(rstatix)
#Calculating p-value for the variation over time
stat.test_time <- alpha_beta_met %>%
  kruskal_test(Shannon ~ Time_tx)
stat.test_time
```

```{r}
#Calculating p-values between treatments by time point
stat.test <- alpha_beta_met %>%
  group_by(Time_tx) %>%
  wilcox_test(Shannon ~ Treatment, alternative = "greater") %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "Time_tx", dodge = 0.8)
stat.test
```
#BOX PLOT WITH GGPUR
```{r}
library(ggpubr)
library(ggsci)
#Boxplot of Shannon diversity between treatment groups over time
shannon_boxplot_bacteria = ggboxplot(alpha_beta_met, x = "Time_tx", y = "Shannon", color = "Treatment", palette = "jama", fill = "Treatment", add = c("jitter"), notch = F, outlier.shape = NA) +
  labs(x = "Time to treatment", y = "Shannon index", colour = "Treatment", fill = "Treatment") + 
  scale_fill_jama(alpha = 0.5) +
  stat_pvalue_manual(stat.test,  label = "p", tip.length = 0) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  annotate("text", x = 2.5, y = 7.5, label = expression(paste("Kruskal-Wallis, ", paste(italic('p'))," = 1.12e-13"))) +
  geom_signif(aes(y = Shannon, x = Time_tx),
              comparisons = list(c("Day -1","Week 1"),c("Week 1","Week 5"),c("Week 5","Week 9") ), 
              test = "wilcox.test", y_position = c(6.7,6.2,6.5), color = "gray35", tip_length = 0.02) +
  theme(axis.title.x = element_blank())

shannon_boxplot_bacteria
```

```{r}
library(rstatix)
#Calculating p-value for the variation over time
stat.test_time <- alpha_beta_met %>%
  kruskal_test(Chao1 ~ Time_tx)
stat.test_time
```

```{r}
#Calculating p-values between treatments by time point
stat.test <- alpha_beta_met %>%
  group_by(Time_tx) %>%
  wilcox_test(Chao1 ~ Treatment, alternative = "greater") %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "Time_tx", dodge = 0.8)

stat.test
```

```{r}
chao1_bacteria_boxplot = ggboxplot(alpha_beta_met, x = "Time_tx", y = "Chao1", color = "Treatment", palette = "jama", fill = "Treatment", add = c("jitter"), notch = F, outlier.shape = NA) +
  labs(x = "Time to treatment", y = "Chao1 index", colour = "Treatment", fill = "Treatment") + 
  scale_fill_jama(alpha = 0.5)  + 
  stat_pvalue_manual(stat.test,  label = "p", tip.length = 0) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  annotate("text", x = 2.5, y = 260, label = expression(paste("Kruskal-Wallis, ", paste(italic('p'))," = 4.21e-11"))) +
  geom_signif(aes(y = Chao1, x = Time_tx),
              comparisons = list(c("Day -1","Week 1"),c("Week 1","Week 5"),c("Week 5","Week 9") ), 
              test = "wilcox.test", y_position = c(230,200,230), color = "gray35", tip_length = 0.02) +
  theme(axis.title.x = element_blank())

chao1_bacteria_boxplot
```

#BETA DIVERSITY
#Bray ordination with densities
```{r}
library(cowplot)
library("scales")
library(ggthemes)

#Permanova value provided by COSMOS-ID
label_perm <- expression(paste("PERMANOVA, ",F ,"= 3.93, ",paste(italic('p')),"=0.001"))

#PCoA of the Bray-Curtis dissimilarity 
pmain_brayF = ggscatter(alpha_beta_met, x = "bray_curtis.X", y = "bray_curtis.Y", color = "Time_tx", shape = "Treatment",ellipse = TRUE, ellipse.level=.7,mean.point = F, mean.point.size = 5, star.plot = F) +
  labs(x = "PCoA 1 (12.95%) ", y = "PCoA 2 (4.41%)", colour = "Time", fill = "Time") + 
  theme(legend.position = "top") +
  scale_fill_colorblind() +   
  scale_color_colorblind() +
  annotate("text", x = -0.3, y = 0.25, label = label_perm, colour = "black") 

#Adding marginal densities along x axis in the PCoA
xdens <- axis_canvas(pmain_brayF, axis = "x") +
  geom_density(data = alpha_beta_met, aes(x = bray_curtis.X, fill = Time_tx), alpha = 0.7, size = 0.2) +
  theme(legend.position = "none") +
  scale_fill_colorblind()

#Adding marginal densities along x axis in the PCoA
ydens <- axis_canvas(pmain_brayF, axis = "y", coord_flip = TRUE) +
  theme(legend.position = "none") +
  geom_density(data = alpha_beta_met, aes(x = bray_curtis.Y, fill = Time_tx), alpha = 0.7, size = 0.2)+
  coord_flip() +
  scale_fill_colorblind()

#Inserting densities in the PCoA plot
p1 <- insert_xaxis_grob(pmain_brayF, xdens, grid::unit(.2, "null"), position = "top")
p2<- insert_yaxis_grob(p1, ydens, grid::unit(.2, "null"), position = "right")

#Ploting the PCoA
bray_plot_bacteria = ggdraw(p2)
bray_plot_bacteria
```

#ABUNDANCE
##Phylum
```{r}
library(reshape2)
#Adding all the abundances at the Phylum level
bacteria_Phylum_ab <- abundance %>% 
  dplyr::select(Phylum, 9:167) %>% 
  group_by(Phylum) %>% 
  summarise_all(sum)

#Transposing table and merging with the metadata
bacteria_Phylum_ab_metadata <- melt(bacteria_Phylum_ab, id = colnames(bacteria_Phylum_ab)[1]) %>% 
  rename(Name = variable, Abundance = value) %>% 
  merge(metadata,by="Name")
```

```{r}
bacteria_Phylum_ab_barplot <- bacteria_Phylum_ab_metadata %>% 
  ggbarplot(x= "Time_tx", y = "Abundance", color = "Phylum", fill = "Phylum", 
            palette = get_palette("rickandmorty",20), 
            ylab = "Relative abundance", 
            legend = "right", 
            position = position_fill()) +
  labs(fill="Phylum", color="Phylum") +
  facet_grid(~Treatment, scales = "free_x", space = "free_x") +
  theme(axis.title.x = element_blank()) +
  guides(color = guide_legend(ncol = 2, bycol = TRUE))

bacteria_Phylum_ab_barplot
```

```{r}
#Renaming less abundant bacteria as a category "Minorities"
bacteria_Phylum_ab_metadata2 <- bacteria_Phylum_ab_metadata %>% 
  mutate(Phylum = recode(Phylum,
                         `Candidatus Fervidibacteria` = "Minorities",
                        Nitrospirae = "Minorities",
                        Balneolaeota = "Minorities",
                        `Candidatus Saccharibacteria` = "Minorities",
                        Aquificae = "Minorities",
                        Gemmatimonadetes ="Minorities",
                        Synergistetes ="Minorities",
                        `Deinococcus-Thermus` = "Minorities",
                        Fusobacteria = "Minorities",
                        Acidobacteria = "Minorities",
                        Tenericutes = "Minorities",
                        Bacteria_u_p = "Minorities",
                        Verrucomicrobia = "Minorities",
                        Chloroflexi = "Minorities"))
```

```{r}
bacteria_Phylum_ab_barplot2 <- bacteria_Phylum_ab_metadata2 %>% 
  ggbarplot(x= "Time_tx", y = "Abundance", color = "Phylum", fill = "Phylum", 
            palette = get_palette("jama",7), 
            ylab = "Relative abundance", 
            legend = "right", 
            position = position_fill()) +
  labs(fill="Phylum", color="Phylum") +
  facet_grid(~Treatment, scales = "free_x", space = "free_x") +
  theme(axis.title.x = element_blank())

bacteria_Phylum_ab_barplot2
```

```{r fig.height=4, fig.width=14}
bacteria_composition_plots <- ggarrange(bray_plot_bacteria,bacteria_Phylum_ab_barplot2,ncol = 2, nrow = 1, align = "hv", widths = c(1,.8), labels = c("A","B"))
bacteria_composition_plots
```

#Composition plots (merging with the plots generated with 3-ARGs_diversity_composition.Rmd)
```{r}
#composition_plots <- ggarrange(bacteria_composition_plots, ARG_composition_plots, nrow = 2)
#composition_plots
```

#Alpha-diversity plots (merging with the plots generated with 3-ARGs_diversity_composition.Rmd)
```{r fig.height=6, fig.width=8}
#alpha_plots <- ggarrange(shannon_boxplot_bacteria, chao1_bacteria_boxplot, shannon_boxplot_ARG, chao1_ARG_boxplot, nrow = 2, ncol = 2, labels = c("A","B","C","D"), common.legend = T)
#alpha_plots
```

#Saving plots in my directory
```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Mastitis_project/figures/composition")
#ggsave(plot=alpha_plots,"alpha_diversity_plots.png", width = 10, height = 8)
#ggsave(plot=composition_plots,"composition_plots.png", width = 17, height = 8)
```