---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
#Physeq object
##Importing tables
```{r echo = FALSE}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Mastitis_project/GitHub/data/")
library(dplyr)
#read in otu table
otu_table = read.csv("feature_table_taxa.csv",sep=",", check.names = F)
otu_table <- otu_table %>% 
  group_by(OTU_ID) %>% 
  dplyr::summarise(across(everything(), sum))

library(tibble)
otu_table <- otu_table %>% remove_rownames %>% column_to_rownames(var = "OTU_ID")
otu_table=as.matrix(otu_table)


#read in taxonomy
taxonomy = read.csv("taxonomy_taxa.csv",sep=",")
taxonomy <-taxonomy %>% distinct(OTU_ID, .keep_all = TRUE)
taxonomy <- taxonomy %>% remove_rownames %>% column_to_rownames(var = "OTU_ID")
taxonomy=as.matrix(taxonomy)

#read in metadata
metadata=read.table("metadata.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 
metadata
```

##Generating a phyloseq object
```{r echo=FALSE}
library("phyloseq")

#import as phyloseq objects
OTU = otu_table(otu_table,taxa_are_rows=TRUE)
TAX = tax_table(taxonomy)
META = sample_data(metadata)

# Sanity checks for consistent OTU names
taxa_names(TAX)
taxa_names(OTU)

# Same sample names
sample_names(OTU)
sample_names(META)

#Final phyloseq output is named as physeq
physeq=phyloseq(OTU,TAX,META)
```

##Filtering samples per time point
```{r}
`%notin%` <- Negate(`%in%`)
physeq_filtered = physeq %>% subset_taxa(Order %notin% "Chloroplast", Family %notin% "Mitochondria")
physeq_day1 <- subset_samples(physeq_filtered, Time_tx%in%"Day -1")
physeq_week1 <- subset_samples(physeq_filtered, Time_tx%in%c("Week 1"))
physeq_week5 <- subset_samples(physeq_filtered, Time_tx%in%c("Week 5"))
physeq_week9 <- subset_samples(physeq_filtered, Time_tx%in%c("Week 9"))
```

#Install DESeq2 if necessary
```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("DESeq2")
BiocManager::install("glmGamPoi")
```

#Test the best DESeq2 fit model
```{r}
library(DESeq2)
library(glmGamPoi)
dds <- phyloseq_to_deseq2(physeq_day1, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="glmGamPoi")
dds_es <- estimateSizeFactors(dds)
par <- estimateDispersions(dds_es, fitType = "parametric")
loc <- estimateDispersions(dds_es, fitType = "local")
me <- estimateDispersions(dds_es, fitType = "mean")
gl <- estimateDispersions(dds_es, fitType = "glmGamPoi")
```

```{r}
plotDispEsts(par, main= "dispEst: parametric")
```

```{r}
plotDispEsts(loc, main= "dispEst: local")
```

```{r}
plotDispEsts(me, main= "dispEst: mean")
```

```{r}
plotDispEsts(gl, main= "dispEst: glmGamPoi")
```

#DAY 1
##Aggregating abundances by taxonomic level
```{r}
day1_Kingdom <- tax_glom(physeq_day1, taxrank = 'Kingdom')
day1_Phylum <- tax_glom(physeq_day1, taxrank = 'Phylum')
day1_Class <- tax_glom(physeq_day1, taxrank = 'Class')
day1_Order <- tax_glom(physeq_day1, taxrank = 'Order')
day1_Family <- tax_glom(physeq_day1, taxrank = 'Family')
day1_Genus <- tax_glom(physeq_day1, taxrank = 'Genus')
day1_Species <- tax_glom(physeq_day1, taxrank = 'Species')

```

##Deseq2 Kingdom
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(day1_Kingdom, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="parametric")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(day1_Kingdom@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
#write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_kingdom_day1_pvalue.csv",sep = ",") 
```

##Deseq2 Phylum
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(day1_Phylum, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(day1_Phylum@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
#write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_phylum_day1.csv",sep = ",") 
```

##Deseq2 Class
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(day1_Class, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(day1_Class@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_class_day1.csv",sep = ",", row.names = F)

```
##Deseq2 Order
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(day1_Order, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(day1_Order@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_order_day1.csv",sep = ",", row.names = F) 
```

##Deseq2 Family
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(day1_Family, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(day1_Family@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_family_day1.csv",sep = ",", row.names = F) 
```
##Deseq2 Genus

```{r}
#Converting phyloseq-format into DESeqDataSet with dispersion estimated by treatment 
dds <- phyloseq_to_deseq2(day1_Genus, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(day1_Genus@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_genus_day1.csv",sep = ",", row.names = F)
```
##Deseq2 Species
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(day1_Species, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(day1_Species@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_species_day1.csv",sep = ",", row.names = F)
```

#WEEK 1
##Aggregating abundances by taxonomic level
```{r}
week1_Kingdom <- tax_glom(physeq_week1, taxrank = 'Kingdom')
week1_Phylum <- tax_glom(physeq_week1, taxrank = 'Phylum')
week1_Class <- tax_glom(physeq_week1, taxrank = 'Class')
week1_Order <- tax_glom(physeq_week1, taxrank = 'Order')
week1_Family <- tax_glom(physeq_week1, taxrank = 'Family')
week1_Genus <- tax_glom(physeq_week1, taxrank = 'Genus')
week1_Species <- tax_glom(physeq_week1, taxrank = 'Species')
```

##Deseq2 Kingdom
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week1_Kingdom, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week1_Kingdom@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_kingdom_week1_pvalue.csv",sep = ",") 
```

##Deseq2 Phylum
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week1_Phylum, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week1_Phylum@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_phylum_week1.csv",sep = ",") 
```

##Deseq2 Class
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week1_Class, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week1_Class@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_class_week1.csv",sep = ",", row.names = F)

```
##Deseq2 Order
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week1_Order, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week1_Order@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_order_week1.csv",sep = ",", row.names = F) 
```

##Deseq2 Family
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week1_Family, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week1_Family@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_family_week1.csv",sep = ",", row.names = F)
```

##Deseq2 Genus
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week1_Genus, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week1_Genus@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_genus_week1.csv",sep = ",", row.names = F)
```

##Deseq2 Species
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week1_Species, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week1_Species@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_species_week1.csv",sep = ",", row.names = F)
```


#WEEK 5
##Aggregating abundances by taxonomic level
```{r}
week5_Kingdom <- tax_glom(physeq_week5, taxrank = 'Kingdom')
week5_Phylum <- tax_glom(physeq_week5, taxrank = 'Phylum')
week5_Class <- tax_glom(physeq_week5, taxrank = 'Class')
week5_Order <- tax_glom(physeq_week5, taxrank = 'Order')
week5_Family <- tax_glom(physeq_week5, taxrank = 'Family')
week5_Genus <- tax_glom(physeq_week5, taxrank = 'Genus')
week5_Species <- tax_glom(physeq_week5, taxrank = 'Species')
```

##Deseq2 Kingdom
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week5_Kingdom, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week5_Kingdom@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
#write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_kingdom_week5_pvalue.csv",sep = ",") 
```

##Deseq2 Phylum
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week5_Phylum, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week5_Phylum@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
#write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_phylum_week5.csv",sep = ",") 
```

##Deseq2 Class
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week5_Class, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week5_Class@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_class_week5.csv",sep = ",", row.names = F)

```
##Deseq2 Order
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week5_Order, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week5_Order@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_order_week5.csv",sep = ",", row.names = F) 
```

##Deseq2 Family
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week5_Family, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week5_Family@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_family_week5.csv",sep = ",", row.names = F) 
```
##Deseq2 Genus
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week5_Genus, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week5_Genus@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_genus_week5.csv",sep = ",", row.names = F)
```

##Deseq2 Species
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week5_Species, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week5_Species@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_species_week5.csv",sep = ",", row.names = F)
```



#WEEK 9
```{r}
week9_Kingdom <- tax_glom(physeq_week9, taxrank = 'Kingdom')
week9_Phylum <- tax_glom(physeq_week9, taxrank = 'Phylum')
week9_Class <- tax_glom(physeq_week9, taxrank = 'Class')
week9_Order <- tax_glom(physeq_week9, taxrank = 'Order')
week9_Family <- tax_glom(physeq_week9, taxrank = 'Family')
week9_Genus <- tax_glom(physeq_week9, taxrank = 'Genus')
week9_Species <- tax_glom(physeq_week9, taxrank = 'Species')
```

##Deseq2 Kingdom
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week9_Kingdom, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="parametric")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week9_Kingdom@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
#write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_kingdom_week9_pvalue.csv",sep = ",") 
```

##Deseq2 Phylum
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week9_Phylum, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week9_Phylum@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_phylum_week9.csv",sep = ",") 
```

##Deseq2 Class
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week9_Class, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week9_Class@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_class_week9.csv",sep = ",", row.names = F)

```
##Deseq2 Order
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week9_Order, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week9_Order@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_order_week9.csv",sep = ",", row.names = F) 
```

##Deseq2 Family
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week9_Family, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week9_Family@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_family_week9.csv",sep = ",", row.names = F) 
```
##Deseq2 Genus
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week9_Genus, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week9_Genus@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_genus_week9.csv",sep = ",", row.names = F)
```
##Deseq2 Species
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week9_Species, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week9_Species@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/taxa/deseq_species_week9.csv",sep = ",", row.names = F)
```