---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

##Importing tables
```{r echo = FALSE}

setwd("~/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Mastitis_project/GitHub/data/")

#read in otu table
otu_table = read.csv("feature_table_genes.csv",sep=",", check.names = F, row.names = 1)
otu_table=as.matrix(otu_table)

#read in taxonomy
taxonomy = read.csv("taxonomy_genes.csv",sep=",", row.names = 1)
taxonomy=as.matrix(taxonomy)

#read in metadata
metadata=read.table("metadata.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 
metadata
```

##Generating a phyloseq object
```{r echo=FALSE}
library("phyloseq")

#import as phyloseq objects
OTU = otu_table(otu_table,taxa_are_rows=TRUE)
TAX = tax_table(taxonomy)
META = sample_data(metadata)

# Sanity checks for consistent OTU names
taxa_names(TAX)
taxa_names(OTU)

# Same sample names
sample_names(OTU)
sample_names(META)

#Final phyloseq output is named as physeq
physeq=phyloseq(OTU,TAX,META)
```

#Filtering samples
```{r}
physeq_day1 <- subset_samples(physeq, Time_tx%in%c("Day -1"))
physeq_week1 <- subset_samples(physeq, Time_tx%in%c("Week 1"))
physeq_week5 <- subset_samples(physeq, Time_tx%in%c("Week 5"))
physeq_week9 <- subset_samples(physeq, Time_tx%in%c("Week 9"))
```

#Install DESeq2 if necessary
```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("DESeq2")
BiocManager::install("glmGamPoi")
```

#Test the best DESeq2 fit model
```{r}
library(DESeq2)
library(glmGamPoi)
dds <- phyloseq_to_deseq2(physeq_day1, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="glmGamPoi")
dds_es <- estimateSizeFactors(dds)
par <- estimateDispersions(dds_es, fitType = "parametric")
loc <- estimateDispersions(dds_es, fitType = "local")
me <- estimateDispersions(dds_es, fitType = "mean")
gl <- estimateDispersions(dds_es, fitType = "glmGamPoi")
```

```{r}
plotDispEsts(par, main= "dispEst: parametric")
```

```{r}
plotDispEsts(me, main= "dispEst: mean")
```

```{r}
plotDispEsts(me, main= "local")
```

```{r}
plotDispEsts(gl, main= "dispEst: glmGamPoi")
```

#DAY 1
##Aggregating abundances by classification level
```{r}
day1_type <- tax_glom(physeq_day1, taxrank = 'Type')
day1_class <- tax_glom(physeq_day1, taxrank = 'Class')
day1_mech <- tax_glom(physeq_day1, taxrank = 'Mechanism')
day1_gene <- tax_glom(physeq_day1, taxrank = 'Gene')
day1_name <- tax_glom(physeq_day1, taxrank = 'Name')
```

##Deseq2 Type
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(day1_type, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(day1_type@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_type_day1.csv",sep = ",") 
```

##Deseq2 Class
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(day1_class, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(day1_class@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_class_day1.csv",sep = ",") 
```

##Deseq2 Mechanism
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(day1_mech, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(day1_mech@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_mechanism_day1.csv",sep = ",") 
```

##Deseq2 Gene
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(day1_gene, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(day1_gene@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_gene_day1.csv",sep = ",") 
```

##Deseq2 Name
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(day1_name, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(day1_name@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_name_day1.csv",sep = ",") 
```


#WEEK 1
##Aggregating abundances by rank
```{r}
week1_type <- tax_glom(physeq_week1, taxrank = 'Type')
week1_class <- tax_glom(physeq_week1, taxrank = 'Class')
week1_mech <- tax_glom(physeq_week1, taxrank = 'Mechanism')
week1_gene <- tax_glom(physeq_week1, taxrank = 'Gene')
week1_name <- tax_glom(physeq_week1, taxrank = 'Name')

```
##Deseq2 Type
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week1_type, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week1_type@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_type_week1.csv",sep = ",") 
```

##Deseq2 Class
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week1_class, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week1_class@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_class_week1.csv",sep = ",") 
```

##Deseq2 Mechanism
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week1_mech, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week1_mech@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_mechanism_week1.csv",sep = ",") 
```

##Deseq2 Gene
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week1_gene, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week1_gene@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_gene_week1.csv",sep = ",") 
```

##Deseq2 Name
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week1_name, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week1_name@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_name_week1.csv",sep = ",") 
```


#WEEK 5
```{r}
week5_type <- tax_glom(physeq_week5, taxrank = 'Type')
week5_class <- tax_glom(physeq_week5, taxrank = 'Class')
week5_mech <- tax_glom(physeq_week5, taxrank = 'Mechanism')
week5_gene <- tax_glom(physeq_week5, taxrank = 'Gene')
week5_name <- tax_glom(physeq_week5, taxrank = 'Name')
```

##Deseq2 Type
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week5_type, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week5_type@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_type_week5.csv",sep = ",") 
```

##Deseq2 Class
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week5_class, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week5_class@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_class_week5.csv",sep = ",") 
```

##Deseq2 Mechanism
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week5_mech, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week5_mech@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_mechanism_week5.csv",sep = ",") 
```

##Deseq2 Gene
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week5_gene, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week5_gene@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_gene_week5.csv",sep = ",") 
```

##Deseq2 Name
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week5_name, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week5_name@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_name_week5.csv",sep = ",") 
```

#WEEK 9
##Aggregating abundances by rank
```{r}
week9_type <- tax_glom(physeq_week9, taxrank = 'Type')
week9_class <- tax_glom(physeq_week9, taxrank = 'Class')
week9_mech <- tax_glom(physeq_week9, taxrank = 'Mechanism')
week9_gene <- tax_glom(physeq_week9, taxrank = 'Gene')
week9_name <- tax_glom(physeq_week9, taxrank = 'Name')
```

##Deseq2 Type
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week9_type, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week9_type@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_type_week9.csv",sep = ",") 
```

##Deseq2 Class
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week9_class, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week9_class@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_class_week9.csv",sep = ",") 
```

##Deseq2 Mechanism
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week9_mech, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week9_mech@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_mechanism_week9.csv",sep = ",") 
```

##Deseq2 Gene
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week9_gene, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week9_gene@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_gene_week9.csv",sep = ",") 
```

##Deseq2 Name
```{r}
#Convering phyloseq-format into DESeqDataSet with dispersions estimated by treatment 
dds <- phyloseq_to_deseq2(week9_name, ~ Treatment)
dds = DESeq(dds, test="Wald", fitType="local")

#Lists the coefficients/comparisons
resultsNames(dds)

#Results table of the desired comparison
res <- results(dds, name="Treatment_Control_vs_Antibiotic") 

#Results with significant p-values
sigtab = res[which(res$pvalue<0.05), ]

#Adding taxonomic names
sigtab <- merge(week9_name@tax_table,sigtab, by = 0)

#Exporting significant results as a csv file (in case you have)
write.table(sigtab, file="~/OneDrive - Michigan State University/Manning_lab/Mastitis_project/tables/diff_abundance/DESeq2/genes/deseq_name_week9.csv",sep = ",") 
```